A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-20 FMCAD2021/3_flatten.tex

The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages e.g. Javascript is more general in the sense that the base can be a number between 2 and 36. Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readability.}.
%Since we focus on the decidability,
%we define a binary version of {\parseInt},
%which takes a binary string and returns a decimal integer number.
For example,
${\parseInt}('0123') = {\parseInt}('123')=10^2+10*2+3 = 123$. 
Note here we use the quotation marks to delimit the strings, 
%Clearly,  our decision procedure given in this paper 
%can be adapted to string constraints with other string to number conversion function without substantial change. 
%Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
%from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.

Formally, the semantics of the $\parseInt$ function is defined as follows. 
%\begin{definition} 
In order to simplify presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$. Note that one can easily extend our approach to allow arbitrary finite alphabet. Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
    for every $w\in \Sigma_{\textit{num}}^+$,
    \begin{itemize}
%        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
        \item  if $w=$`$i$' for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}$$($`$i$'$)=i$;
        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}($`$i$'$)$.
    \end{itemize} 
%\end{definition}
Note that $\parseInt$ is undefined with $\varepsilon$ as the input.

@@

A \emph{flat domain restriction} for a string constraint $\varphi$ is a function $\flatfun_\varphi$ that maps each string variable $\svarx \in \svar(\varphi)$ to a flat language $(w_{\svarx,1})^* \cdots (w_{\svarx, k_\svarx})^*$, where $w_{\svarx, i} \in \Sigma_{\textit{num}}^+$ for every $i \in [k_\svarx]$. 
%
The flattened semantics of $\phi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.

The flattening of $\varphi \in \strint$ under a flat domain restriction $\flatfun_\varphi$ is a {\paexp} formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes its flattened semantics.
%
More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and Parikh variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, called \emph{flattening variables}, plus some other auxiliary variables, such that 
%
$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
%
The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \Nat$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that 
\begin{itemize}
\item  
for every $ \svarx \in \svar(\varphi)$, $I_s(\svarx) = w_{\svarx,1}^{I_e(\#_{\svarx,1})} \ldots  w_{\svarx,k_\svarx}^{I_e(\#_{\svarx,k_\svarx})}$, 
%
\item for every $ \ivarx \in \ivar(\varphi)$, $I_i(\ivarx) = I_e(\ivarx)$.
\end{itemize}

@@

Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$, we define $\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) = \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows: 
\begin{itemize}
\item for $i = k_\svarx$, 
$$\iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
%
\item for $i \in [k_\svarx-1]$, 
%
$$\iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
%
and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
\end{itemize}
Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.

Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure.
\begin{enumerate}
\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by 
 $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
%
\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows: 
\begin{itemize}
\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} |w_{\alpha_m, j}| \#_{\alpha_m, j}$,
%
\item for $i \in [m-1]$, if $\alpha_i \in \Sigma_{\textit{num}}$, then $\iterm_{t, i} = \alpha_i 10^{\ell_{t, i+1}} + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1}+1$, otherwise, $\iterm_{t, i} = 
\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j}| \#_{\alpha_i, j}$.
\end{itemize}
%
\item Let $\iterm'_1 \op \iterm'_2$ be the formula resulting from the aforementioned replacements. (Note that strictly speaking, $\iterm'_1 \op \iterm'_2$ is not a $\paexp$ formula since it contains divisions.) Let $N$ be the least common multiplier of the denominators of $\iterm'_1$ and $\iterm'_2$.  Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained by multiplying the both sides of $\iterm'_1 \op \iterm'_2$ with $N$,  so that the division operator disappears. 
\end{enumerate}

\begin{example}
Suppose $\parseInt(\svarx) = 2\ivarx$ is an atomic constraint and $\flatfun_\varphi(\svarx) = 1^*2^*$. Then 
\[
\small
\begin{array}{l l}
& \flatten_{\flatfun_\varphi}(\parseInt(\svarx)  =  2\ivarx)  \\
\Def & 1 \frac{10^{\#_{\svarx,1}}-1}{10-1}10^{\#_{\svarx,2}}  + 2 \frac{10^{\#_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2 (10^{2\#_{\svarx,2}}-1) = 18\ivarx \\
\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} +  10^{\#_{\svarx,2}} = 18\ivarx + 2.
\end{array}
\]
%Take $n={\parseInt}((11)^a(10)^b)$ for example.
%\begin{align}
%    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
%    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
%    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
%\end{align}
\end{example}

>>>>

The function  {\parseInt} takes a decimal string as the input and returns the integer represented by the string\footnote{The {\parseInt} function in scripting languages e.g. Javascript is more general in the sense that the base can be a number between 2 and 36. Although our approach works for arbitrary positive bases, we choose to focus on the base 10 in this paper, for readability.}.
%Since we focus on the decidability,
%we define a binary version of {\parseInt},
%which takes a binary string and returns a decimal integer number.
For example,
${\parseInt}($`$0123$'$) = {\parseInt}($`$123$'$)=10^2+10*2+3 = 123$. 
Note here we use the quotation marks to delimit the strings.
%Clearly,  our decision procedure given in this paper 
%can be adapted to string constraints with other string to number conversion function without substantial change. 
%Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
%from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.

Formally, the semantics of the $\parseInt$ function is defined as follows. 
%\begin{definition} 
In order to simplify the presentation, we assume all string variables ranging over numerical symbols $\Sigma_{\textit{num}}=\{0,1, \ldots, 9\}$. Note that one can easily extend our approach to allow arbitrary finite alphabet. Then ${\parseInt}: \Sigma_{\textit{num}}^+ \mapsto \Nat$ is recursively defined by
    for every $w\in \Sigma_{\textit{num}}^+$,
    \begin{itemize}
%        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
        \item  if $w=$`$i$' for $i \in \Sigma_{\textit{num}}$, then ${\parseInt}$$($`$i$'$)=i$;
        \item for $w = w_1$`$i$' for $i \in \Sigma_{\textit{num}}$ with $|w_1| \ge 1$, 
        ${\parseInt}(w) = 10*{\parseInt}(w_1)+{\parseInt}($`$i$'$)$.
    \end{itemize} 
%\end{definition}
Note that $\parseInt$ is undefined with $\varepsilon$ as the input.

@@

A \emph{flat domain restriction} for a string constraint $\varphi$ is a function $\flatfun_\varphi$ that maps each string variable $\svarx \in \svar(\varphi)$ to a flat language $(w_{\svarx,1})^* \cdots (w_{\svarx, k_\svarx})^*$, where $w_{\svarx, i} \in \Sigma_{\textit{num}}^+$ for every $i \in [k_\svarx]$. 
%
The flattened semantics of $\varphi \in \strint$ is defined as $\llbracket \varphi \rrbracket_{\flatfun_\varphi} = \{(I_s, I_i) \in \llbracket \varphi  \rrbracket \mid \forall \svarx \in \svar(\varphi).\ I_s(x) \in  \flatfun_\varphi(\svarx)\}$.

The flattening of $\varphi \in \strint$ under a flat domain restriction $\flatfun_\varphi$ is a {\paexp} formula, denoted by $\flatten_{\flatfun_\varphi}(\varphi)$, that encodes its flattened semantics.
%
More concretely, $\flatten_{\flatfun_\varphi}(\varphi)$ is a formula over the integer variables $\ivar(\varphi)$,  and flattening variables $\pvar_{\flatfun_\varphi}(\varphi) = \bigcup_{\svarx \in \svar(\varphi)} \pvar_{\flatfun_\varphi}(\svarx)$, where $\pvar_{\flatfun_\varphi}(\svarx) = \{\#_{\svarx,i} \mid i \in [k_\svarx]\}$, plus some other auxiliary variables, such that 
%
$$\llbracket \varphi \rrbracket_{\flatfun_\varphi} =\decode_{\flatfun_\varphi}(\llbracket \flatten_{\flatfun_\varphi}(\varphi) \rrbracket |_{\ivar(\phi) \cup \pvar_{\flatfun_\varphi}(\varphi)})$$
%
The decoding function above decodes an interpretation of integer and flattening variables $I_e: \ivar(\varphi) \cup \pvar_{\flatfun_\varphi}(\varphi) \rightarrow \Nat$ as a set $\decode_{\flatfun_\varphi}(I_e)$ of interpretations of the $\varphi$'s integer and string variables $(I_s, I_i)$ with $I_s: \svar(\varphi) \rightarrow \Sigma^*$ and $I_i: \ivar(\phi) \rightarrow \Nat$ such that 
\begin{itemize}
\item  
for every $ \svarx \in \svar(\varphi)$, $I_s(\svarx) = w_{\svarx,1}^{I_e(\#_{\svarx,1})} \ldots  w_{\svarx,k_\svarx}^{I_e(\#_{\svarx,k_\svarx})}$, 
%
\item for every $ \ivarx \in \ivar(\varphi)$, $I_i(\ivarx) = I_e(\ivarx)$.
\end{itemize}

@@

For simplicity, we assume that each occurrence of $\parseInt$ (resp. $\len(t)$) in $\iterm_1 \op \iterm_2$ is of the form $\parseInt(\svarx)$ (resp. $\len(\svarx)$) for a string variable $\svarx$. (Otherwise, we can always introduce a fresh string variable $\svarx'$ for $\parseInt(\sterm)$ or $\len(\sterm)$ and add the constraint $\svarx' = \sterm$.)
Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by replacing $\parseInt(\svarx)$ with $\flatten_{\flatfun_\varphi}(\parseInt(\svarx))$ and $\len(\svarx)$ with $\flatten_{\flatfun_\varphi}(\len(\svarx))$, where 
\begin{itemize}
\item 
$\flatten_{\flatfun_\varphi}(\parseInt(\svarx)) \Def \iterm_{\svarx,1}$  such that $(\iterm_{\svarx, i})_{i \in [k_\svarx]}$ and $(\ell_{\svarx, i})_{i \in [k_\svarx]}$ are inductively defined as follows: 
\begin{itemize}
\item for $i = k_\svarx$, 
$$ \hspace{-4mm} \iterm_{\svarx, i} = \frac{\parseInt(w_{\svarx, k_\svarx}) (10^{|w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx} } -1 )}{(10^{|w_{\svarx, k_\svarx}|} -1 )}$$ 
and $\ell_{\svarx, i} = |w_{\svarx,k_\svarx}| \#_{\svarx, k_\svarx}$,
%
\item for $i \in [k_\svarx-1]$, 
%
$$ \hspace{-1.2cm} \iterm_{\svarx, i} =  \frac{\parseInt(w_{\svarx, i}) (10^{|w_{\svarx, i} | \#_{\svarx, i} } -1 ) 10^{\ell_{\svarx, i+1}}} {(10^{|w_{\svarx, i}|} -1 )} + \iterm_{\svarx, i+1}$$
%
and $\ell_{\svarx, i} = |w_{\svarx, i} | \#_{\svarx, i} + \ell_{\svarx, i+1}$.
%
\end{itemize}
Notice that here $|w_{\svarx, \textunderscore } |$ and $\ell_{\svarx, \textunderscore}$ are constants while $\#_{\svarx, \textunderscore}$ are (flattening) variables.
%
\item $\flatten_{\flatfun_\varphi}(\len(\svarx)) \Def \sum \limits_{i \in [k_{\svarx}]} |w_{\svarx, i} | \#_{\svarx, i}$. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
Before presenting the construction of $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$, for every $\svarx \in \svar(\varphi)$,

Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained from $\iterm_1 \op \iterm_2$ by the following procedure.
\begin{enumerate}
\item Replace each integer term $\len(t)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by 
 $\sum \limits_{i \in [m]} \theta_i$, where for every $i \in [m]$, $\theta_i = 1$ if $\alpha_i \in \Sigma_{\textit{num}}$ and $ \theta_i = \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j} | \#_{\alpha_i, j}$ otherwise.
%
\item  Replace each integer term $\parseInt(\sterm)$ such that $t = \alpha_1 \ldots \alpha_m$ with $\alpha_i \in \Sigma_{\textit{num}} \cup \svar(\varphi)$ for every $i \in [m]$, by $\iterm_{\parseInt(\sterm)}$, where $\iterm_{\parseInt(\sterm)} = \iterm_{t,1}$ such that $(\iterm_{t,i})_{i \in [m]}$ and $(\ell_{t, i})_{i \in [m]}$ are inductively defined as follows: 
\begin{itemize}
\item if $\alpha_m \in \Sigma_{\textit{num}}$, then $\iterm_{t, m} = \alpha_m$ and $\ell_{t, m} = 1$, otherwise, $\iterm_{t, m} = \flatten_{\flatfun_\varphi}(\parseInt(\alpha_m))$ and $\ell_{t, m} = \sum \limits_{j \in [k_{\alpha_m}]} |w_{\alpha_m, j}| \#_{\alpha_m, j}$,
%
\item for $i \in [m-1]$, if $\alpha_i \in \Sigma_{\textit{num}}$, then $\iterm_{t, i} = \alpha_i 10^{\ell_{t, i+1}} + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1}+1$, otherwise, $\iterm_{t, i} = 
\flatten_{\flatfun_\varphi}(\parseInt(\alpha_i))10^{\ell_{t, i+1}}  + \iterm_{t, i+1}$ and $\ell_{t, i} = \ell_{t, i+1} + \sum \limits_{j \in [k_{\alpha_i}]} |w_{\alpha_i, j}| \#_{\alpha_i, j}$.
\end{itemize}
%
\item Let $\iterm'_1 \op \iterm'_2$ be the formula resulting from the aforementioned replacements. (Note that strictly speaking, $\iterm'_1 \op \iterm'_2$ is not a $\paexp$ formula since it contains divisions.) Let $N$ be the least common multiplier of the denominators of $\iterm'_1$ and $\iterm'_2$.  Then $\flatten_{\flatfun_\varphi}(\iterm_1 \op \iterm_2)$ is obtained by multiplying the both sides of $\iterm'_1 \op \iterm'_2$ with $N$,  so that the division operator disappears. 
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{example}
Suppose $\parseInt(\svarx) = 2\ivarx$ is an atomic constraint and $\flatfun_\varphi(\svarx) = 1^*2^*$. Then 
\[
\small
\begin{array}{l l}
& \flatten_{\flatfun_\varphi}(\parseInt(\svarx)  =  2\ivarx)  \\
\Def & 1 \frac{10^{\#_{\svarx,1}}-1}{10-1}10^{\#_{\svarx,2}}  + 2 \frac{10^{\#_{\svarx,2}}-1}{10-1} = 2\ivarx   \\
\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} - 10^{\#_{\svarx,2}}  + 2*10^{2\#_{\svarx,2}}- 2 = 18\ivarx \\
\equiv & 10^{\#_{\svarx,1}+\#_{\svarx,2}} +  10^{\#_{\svarx,2}} = 18\ivarx + 2.
\end{array}
\]
%Take $n={\parseInt}((11)^a(10)^b)$ for example.
%\begin{align}
%    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
%    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
%    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
%\end{align}
\end{example}

### 2021-05-20 FMCAD2021/5_optimization.tex

Let $\varphi$ be a quantifier-free {\paexp} formula in the remainder of this section. Moreover, we assume that $\varphi$ is normalized since the optimizations presented in the sequel are for normalized formulas. Furthermore, for technical convenience, we assume that all the inequality atomic formulas are of the form 
%
\[\sum_{j=1}^{n} a_j 10^{\ivarx_{j}} + \sum_{k=1}^{n}b_k \ivarx_{k} \le c\]
%
where $c$ is an integer constant. (Implicitly, we assume that there are no free variables and all the variables are existentially quantified.)

@@

In Section~\ref{sec-dec}, the decision procedure of {\paexp} requires the elimination of both exponential and linear occurrences of variables. Considering the fact that {\pa} formulas can be solved efficiently by the state-of-the-art solvers, e.g. CVC4 and Z3, one natural idea is to try to eliminate only the exponential occurrences, but not the linear occurrences, of variables, and obtain the {\pa} formulas in the end, which can then be solved by the state-of-the-art solvers.

@@

Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, namely, 
$$
\begin{array}{l}
\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
\hspace{4mm} 
\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c,
\end{array}
$$
then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$. In this case, the quantifier elimination algorithm of {\pa} has to be applied to eliminate $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, so that later on, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-1)}$, which requires that $\ivarx_{\sigma(i-1)}$ is the maximum variable in the left-hand side of the inequality.

@@

We also propose some additional optimizations, which are omitted due to the page limit and can be found in Appendix~\ref{app-opt}.

>>>>

Let $\varphi$ be a quantifier-free {\paexp} formula in the remainder of this section. Moreover, we assume that $\varphi$ is normalized since the optimizations presented in the sequel are for normalized formulas. Furthermore, for technical convenience, we assume that all the inequality atomic formulas are of the form 
%
$\sum_{j=1}^{n} a_j 10^{\ivarx_{j}} + \sum_{k=1}^{n}b_k \ivarx_{k} \le c$, 
%
where $c$ is an integer constant. (Implicitly, we assume that there are no free variables and all the variables are existentially quantified.)

In the sequel, we present two major optimizations and omit some additional optimizations due to the page limit, which can be found in Appendix~\ref{app-opt}.

@@

The decision procedure of {\paexp} in Section~\ref{sec-dec} requires the elimination of both exponential and linear occurrences of variables. Considering the fact that {\pa} formulas can be solved efficiently by the state-of-the-art solvers, e.g. CVC4 and Z3, one natural idea is to try to only eliminate the exponential occurrences, but not the linear occurrences, of variables, and obtain the {\pa} formulas in the end, which can then be solved by the state-of-the-art solvers.

@@

Nevertheless, the side condition of Lemma~\ref{lem:exp-ineq}, namely $a_i \neq 0$, undermines the aforementioned natural idea. If $a_i = 0$, but $b_i \neq 0$, 
%namely, 
%$$
%\begin{array}{l}
%\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) \Def  \\
%\hspace{4mm} 
%\sum_{j=1}^{i-1} a_j 10^{\ivarx_{\sigma(j)}} + \sum_{k=1}^{i}b_k \ivarx_{\sigma(k)} \le c,
%\end{array}
%$$
then we are unable to utilize Lemma~\ref{lem:exp-ineq} to eliminate the linear occurrences of $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$. In this case, the quantifier elimination algorithm of {\pa} has to be applied to eliminate $\ivarx_{\sigma(i)}$ from $\tau(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})$, so that later on, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-1)}$, which requires that $\ivarx_{\sigma(i-1)}$ is the maximum variable in the left-hand side of the inequality.

@@

%We also propose some additional optimizations, which are omitted due to the page limit and can be found in Appendix~\ref{app-opt}.
